How Slow do my Transient Type Checks Go?

### 2019-08-17 paper.tex
Aded a sentence about not finding other patterns

To verify this hypothesis 
% As an alternative, less labour intensive approach to identifying
% potentially influential type annotations,
we measured for each type annotation one additional
configuration, with only that single type annotation present.
Our aim was to compare the overhead of
each type annotation in isolation against the no-typecheck baseline.

Figure \ref{f:pattern} shows the results of these experiments.
% , together with patterns we manually identified.
It shows a pair of graphs for a selection of ten benchmarks: a scatterplot and a column graph.

The column graphs, right of the corresponding scatter plot, show the execution time of configurations with only a single type annotation, the x-axis indicates the index of this annotation, thus the first column represents the first annotation, and the last column represents the last one (in the order they appear in the source code). The y-axes for the column graphs are the same as the associated scatter graphs. However, though the right y-axes for the column graphs appear to be identical to the scatter plot ones, they are relative to performance results collected on different days: this was to ensure that the performance of the test computer had not changed significantly from when the typing lattice experiments were run.

@@

For each benchmark we highlighted one or two type annotations that appear to show a pattern for the typing lattice performance, or in the case of SpectralNorm and Json had higher than usual columns. The identified types are those represented by the red and blue columns. The scatter plots are colour-coded accordingly: a red or blue circle represents configurations with the given type annotation present, but not both, purple circles represent configurations with both type annotations present, and grey circles represent those with neither.

\here
As can be seen, most of the time the patterns we identified in the typing lattice graphs correspond to outliers in the single type annotation column graphs. For the Snake, Towers and CD benchmarks, there is clearly a pattern where an individual type annotation (highlighted in red) appears to be the cause of the slowest/upper half of the typing lattice graphs. For Permute there is actually a significant performance \emph{increase}, indicated by the lower half of the lattice being highlighted in red. DeltaBlue and Go also show a slight increase in performance, but here this is caused by two type annotations (in red and blue); this effect does not appear to be cumulative, however: the purple dots are about the same height as the red and blue ones, so \emph{either} annotation is sufficient for the full benefit.

>>>>

To verify this hypothesis 
% As an alternative, less labour intensive approach to identifying
% potentially influential type annotations,
for each type annotation we measured one additional
configuration, with only that single type annotation present.
We did this to compare the overhead of
each type annotation in isolation against the no-typecheck baseline.

Figure \ref{f:pattern} shows the results of these experiments.
% , together with patterns we manually identified.
It shows a pair of graphs for a selection of ten benchmarks: it's associated typing lattice scatter plot and a column graph showing the results of these single type annotation experiments.

The column graphs, right of the corresponding scatter plot, shows the execution time of configurations with only a single type annotation, the x-axis indicates the index of this annotation, thus the first column represents the first annotation, and the last column represents the last one (in the order they appear in the source code). The y-axes for the column graphs are the same as the associated scatter graphs. However, though the right y-axes for the column graphs appear to be identical to the scatter plot ones, they are relative to performance results collected on different days: this was to ensure that the performance of the test computer had not changed significantly from when the typing lattice experiments were run.

@@

For each benchmark we highlighted one or two type annotations that appear to show a pattern for the typing lattice performance scatter plots, or in the case of SpectralNorm and Json had higher than usual columns. The identified types are those represented by the red and blue columns. The scatter plots are colour-coded accordingly: a red or blue circle represents configurations with the given type annotation present, but not both, purple circles represent configurations with both type annotations present, and grey circles represent those with neither. Though we exhaustively inspected such colour-coded scatter plot for all 1,775 type annotations across all 21 benchmarks, the only patterns we noticed are those shown in \ref{f:pattern}.

As can be seen, most of the time the patterns we found in the typing lattice graphs correspond to outliers in the single type annotation column graphs. For the Snake, Towers and CD benchmarks, there is clearly a pattern where an individual type annotation (highlighted in red) appears to be the cause of the slowest/upper half of the typing lattice graphs. For Permute there is actually a significant performance \emph{increase}, indicated by the lower half of the lattice being highlighted in red. DeltaBlue and Go also show a slight increase in performance, but here this is caused by two type annotations (in red and blue); this effect does not appear to be cumulative, however: the purple dots are about the same height as the red and blue ones, so \emph{either} annotation is sufficient for the full benefit.
